%% HOW TO USE THIS TEMPLATE: %%
%% Type your solution to each problem part within
%% the \begin{solution} \end{solution} environment immediately
%% following it.  Use any of the macros or notation from the
%% header.tex that you need, or use your own (but try to stay
%% consistent with the notation used in the problem).
%%
%% If you have problems compiling this file, you may lack the
%% Header.tex file (available on the course box page), or your system
%% may lack some LaTeX packages.  The "exam" package (required) is
%% available at:
%%
%% http://mirror.ctan.org/macros/latex/contrib/exam/exam.cls
%%
%% Other packages can be found at ctan.org, or you may just comment
%% them out (only the exam and ams* packages are absolutely required).


% The "answers" option causes the solutions to be printed.
% The pdf for this homework that's on blackboard was compiled without the
% answers flag for compactness. You need to include it.
%\documentclass[11pt,addpoints]{exam}
\documentclass[11pt,addpoints,answers]{exam}

% required macros -- get latest hwheader.tex file from blackboard if compiling
\input{hwheader}

\newcommand{\TSAT}{\prob{3-SAT} }
\newcommand{\ILP}{\prob{Integer\ Linear\ Programming} }
\newcommand{\IS}{\prob{Independent\ Set} }
\newcommand{\VC}{\prob{Vertex\ Cover} }
\newcommand{\CQ}{\prob{Clique} }
\newcommand{\HC}{\prob{Hamiltonian\ Cycle} }
\newcommand{\HP}{\prob{Hamiltonian\ Path} }
\newcommand{\KS}{\prob{Knapsack} }
\newcommand{\TSP}{\prob{Traveling\ Salesman\ Problem} }
\newcommand{\SSM}{\prob{Subset\ Sum} }
\newcommand{\SCov}{\prob{Set\ Cover} }

\hwheader{Homework 7}
{Are Oelsner}
{Spring 2017, Bhakta}
{CS 315, Algorithms}
{Due: 3:00pm Friday, Dec 8}

% VARIABLES

\begin{document}

\pagestyle{head}                % put header on every page

\noindent 
{\bf Directions for Homework 7}: You may assume from class that the following
problems (both the search and decision versions) are NP-complete when doing
your reductions. We will (eventually) cover the reasons that all these
problems are NP-complete in class.
\begin{itemize}
  \item \SAT: Given a boolean formula $F$ in CNF form with $n$ variables and
    $m$ clauses, determine if there is an assignment of the $n$ variables that
    satisfies $F$.
  \item \TSAT: A special case of \SAT where every clause has at most 3
    literals.
  \item \ILP: Given a set of $n$ variables and $m$ linear constraints, and a
    linear objective function, find an \emph{integer} solution to the
    constraints that optimizes the objective.
  \item \IS: Given graph $G$ and $k\in\Z^+$, determine if $G$ has an
    indep. set of size $\geq k$.
  \item \VC: Given graph $G$ and $k\in\Z^+$, determine if $G$ has a vertex
    cover of size $\leq k.$
  \item \CQ: Given graph $G$ and $k\in\Z^+$, determine if $G$ has a clique
    of size $\geq k$.
  \item \SSM: Given a list of integers $A$ and a target $T$, determine if
    some sublist of $A$ adds up to $T$.
  \item \KS: Given a list of $n$ items with weights $w_i$ and values $v_i$
    a maximum capacity $C$ and a target value $V$, determine if some subset of
    of the items have a total weight $\leq C$ and a total value $\geq V$.
  \item \HP(Rudrata):  Given graph $G$, determine if $G$ has a path that
    visits every vertex exactly once.
  \item \HC(Rudrata):  Given graph $G$, determine if $G$ has a cycle that
    visits every vertex exactly once.
\end{itemize}

\begin{questions}
  \question[20]
  We define the problem $\CQ3(G,k)$.  You are tasked with determining if $G$
  has a clique of size $k$, but only on speical inputs where the max degree of
  any vertex of $G$ is $3$. In otherwords, this is a special case of the
  $\CQ(G,k)$ problem, but we are restricted to graphs in which every vertex
  has degree at most 3.
  \begin{parts}
    \part Prove that $\CQ3$ is in NP.
    \begin{solution}
      We could easily check a provided solution to $\CQ3$ in linear time as the
      maximum size clique would be 4 vertices, so if the returned clique is
      greater than 4 we could automatically return false. If the clique is 4 or
      fewer vertices we could check each of those vertices to make sure that
      they are connected to each of the other vertices. In the worst case, when
      you have a clique of size 4, you would have to check a total of 6 edges.
      Thus, $\CQ3$ is in NP. 
    \end{solution}

    \part I claim that $\CQ3$ is NP-complete.
    \begin{proof}
      We showed that $\CQ3$ is in NP in part a.

      We show a reduction from $\CQ3$ to \CQ, a known NP-hard problem.
      Clique is a generalizaion of $\CQ3$, since $\CQ3$ consists of the
      special case of \CQ where the input graph has max degree $3$.
      Therefore we can solve $\CQ3(G,k)$ by returning $\CQ(G,k)$ on exactly the
      input to $\CQ3$ unchanged.
      
      This proves the correctness of the reduction and, therefore the
      NP-completeness of Clique-3.
    \end{proof}

    What is wrong with the above proof?

    \begin{solution}
      This 'proof' attempts to show that $\CQ3$ is NP-complete by reducing it to
      $\CQ$, however, this does not prove that $\CQ3$ is NP-complete. To show
      that a problem is NP-complete you need to show that an NP-complete problem
      reduces to it, not the other way around. 
    \end{solution}

    \part Here's a true fact that you can use without proof.

    \begin{theorem}
      $\IS3(G,k)$, the special case of $\IS(G,k)$ restricted to graphs of
      degree at most 3, is NP-complete.
    \end{theorem}

    Here's another fact that we learned in lecture when showing that \CQ
    was NP-complete.
    \begin{theorem}
      A subset $X \subseteq V$ is an independent set in $G$ if and only if
      $X$ is a clique in the ``opposite'' graph $G' = (V,E' = {v \choose 2}
      \setminus E)$.
      Therefore $G$ has a independent set of size $\geq k$ if and only $G'$
      has a clique of size $\geq k$.
    \end{theorem}

    Now lets try again.

    Claim: $\CQ3$ is NP-complete.
    
    \begin{proof}
      $\CQ3$ is in NP as proved in part a.

      We present a reduction from $\IS3$ to $\CQ3$. 
      
      To solve $\IS3(G,k)$, we first create $G'$ by switching all the edges of
      $G$, and returning $\CQ3(G',k)$.

      Proof of correctness:
      By the second fact, $G$ has an independent
      set of size $k$ if and only if $G'$ has a clique of size $k$. This
      proves the correctness of the reduction.
      
      By the first fact, $\IS3$ is NP-complete, and therefore $\CQ3$ is
      NP-complete.
    \end{proof}

    What's wrong with the proof this time?
    \begin{solution}
      When you switch all of the edges to construct $G'$ the graph property of
      each vertex having a degree of at most three would likely not be
      maintained. 
    \end{solution}

    \part Describe any poly-time algorithm for $\CQ3$. Hint: How big can a
    clique get?
    \begin{solution}
      A clique in this problem can't be made up of more than four vertices so if
      k is greater than four return false. 
      The algorithm could first remove any vertices from the graph with degree
      less than k-1. After those vertices and associated edges have been removed
      you could repeat the process on the leftover vertices with their new
      degrees, and repeat that until there are no vertices with degree less than
      k-1 left. Then, for each of the remaining vertices check the number of
      edges between each of its neighbors, meaning that for a vertex A with
      neighbors B, C, and D, check the number of edges between B, C, and D. For
      a 4 clique, return true if that number of edges is equal to 3. For a 3
      clique return true if the number of edges is greater than or equal to 1.
      For a 2 clique return true if it has any neighbors at all. For a 1 clique
      return true if the graph is not empty. If none of the vertices in the
      final graph return true, return false. 
    \end{solution}

  \end{parts}

  \question[40] Generalizations:
  Show that each of the following problems are NP-hard. (Don't worry about
  showing that these problems are in NP). Each reduction should be ``simple'' -
  each problem below is either a direct generalization of a known
  NP-hard problem, or solves a known NP-hard problem with a relatively simple
  transformation (like from Independent Set to Clique).
  \begin{parts}
    \part (EXAMPLE) \prob{Weighted\ Independent\ Set}: Given a graph $G =
    (V,E)$ with integer weights on the vertices $V$, and number $k$, determine
    if $G$ has an independent set of total weight $\geq k$.\\ Prove that this
    problem is NP-complete.
    \begin{solution}
      \begin{itemize}
        \item First we show this is in NP. Say we are given the inputs $G,K$ as
          well as an independent set with total weight $\geq k$. Our verifier
          will check that
          \begin{itemize}
            \item The given vertices have total weight $\geq k$ in at most
              $O(n)$ time.
            \item That this is a valid independent set, by checking each edge
              and making sure that both endpoints are never in the set. This
              will take $O(E)$ time.
            \item Accept if both of the above are True
          \end{itemize}
          This will accept iff the certificate is a valid independent set of
          total weight $\geq k$, which exists iff there is an independent set of
          total weight $\geq k$.
        \item 
          We show that this problem is a generalization of \IS.

          Our reduction: Given an instance $(G,k)$ of $\IS$, create a weighted
          version of $G$ where every vertex has weight $1$. Call this graph
          $G'$. Now, return $\prob{Weighted\ Independent\ Set}(G',k)$.
          \begin{proof}
            By construction, any set of $k$ vertices in $G$ has total weight $k$
            in $G'$ and vice versa. Thus an independent set of $k$ vertices
            exists in $G$ iff that independent set has total weight $k$ in $G'$.
          \end{proof}
      \end{itemize}

    \end{solution}



    \part (EXAMPLE) \prob{SAT_{\leq k}}:
    Given a SAT instance with $n$ variables and $m$ clauses, and an integer $k
    \leq n,$ decide if there is a satisfying assignment in which \emph{at most}
    $k$ of the $n$ variables are assigned TRUE.\\
    Prove that this problem is NP-complete.
    \begin{solution}
      \begin{itemize}
        \item First we show this is in NP. Say we are given the input formula
          $f$, the integer $k$, and a satisfying assignment that sets at most
          $k$ variables ot true. Our verifier will check that
          \begin{itemize}
            \item At most $k$ of the variables are set to true. We can check
              this in $O(n)$ time.
            \item That this is a satisfying assignment by evaluating the
              formula. Each clause takes $O(n)$ time to check as there are at
              most $n$ variables per clause. There are $m$ total clauses, so
              this takes a total of in $O(mn)$ time.
            \item Accept if both of the above are True
          \end{itemize}
          This will accept iff there is a valid assignment
        \item 
          We show that this problem is a generalization of \SAT.

          Our reduction: On an input $F$ to \SAT, we return $\prob{SAT_{\leq
          n}}(F)$.
          
          \begin{proof}
          Since the restriction ``at most $n$ of the variables are set to True''
          applies to any assignment, the problem $\SAT_{\leq n}$ is exactly the
          problem of $\SAT(F)$, and $\SAT$ is a special case of $\SAT_{\leq n}$.
          \end{proof}
      \end{itemize}
    \end{solution}

    \part Given a set of $n$ elements $E$, and a family of
    $m$ subsets $S_i \subseteq E$, and ``budget'' $b$, determine if there are
    some $b$ of these subsets $\set{S_i}$ such that their union is $E$.
    \begin{solution}
      I will show that this problem is a generalization of Vertex Cover.\\

      Reduction: Vertex Cover has inputs graph G and integer k. From G construct
      a set E such that it contains all of the edges of G. For each vertex v in
      G construct a subset $S_v$ of E such that each subset $S_v$ contains every
      edge connected to v. Run the algorithm on following input with values of k
      from 1 to k ($E, S, k$), returning true if the union is E, or false after
      unsuccessfully trying with b = k. 


      \textbf{\underline{Wrong Answer}: Not related to question but this shows
      reduction from Set Cover to ILP}\\
      I will show that Integer Linear Programming is a generalization of Set
      Cover.\\
      I would construct a matrix A such that the n rows represent each of the
      elements in E and the m columns represent each of the subsets. If the row
      element is present in the column subset it is set to 1, otherwise 0. \\
      The optimization problem would attempt to solve the equation below such
      that each of the n indices of x can have a value of either 1 or 0,
      representing if that subset is included or not. b is a vector of size m
      with every index equal to 1 showing that each element needs to be included
      at least once.\\
      $Ax \geq b$\\
      The end goal is to find an assignment of 1 and 0 values to the indices of
      x such that the sum of those values is equal to the budget b. \\

      Reduction: Given an instance (E[n], S[m], b) of Set Cover, create the
      matrix A described above and optimize the equation above with the
      constraints discussed, so that $\sum_{i=0}^{n-1}x_i = budget\ b$. \\
      $A_i =$ 1 if element A is in subset i, 0 if not. \\
      For each element $A \in E$\\
      $A_{i,0}x_0+A_{i,1}x_1+...+A_{i,n}x_{n} \geq 1$\\
      $x_i \in Z^+,  0 \leq x_i \leq 1$\\

      If you are attempting to find at most b subsets such that their union is E
      you would optimize $\sum_{i=0}^{n-1}x_i \leq b$
    \end{solution}

    \part \prob{Longest\ s-t\ Path }:
    Given a graph $G$, two vertices $s$ and $t$ in $V$, and an integer $k$
    determine if $G$ has a simple path (no repeat vertices) that uses $k$ or
    more edges from $s$ to $t$.
    \begin{solution}
      I will show that this problem is a generalization of Hamiltonian Paths.\\
      A Hamiltonian path with input G returns true if there is a simple path
      that goes through every vertex in V. Such a path would have to have a
      length of V-1. 

      Reduction: Run the Longest Path algorithm on every pair of vertices in V,
      returning true if there is a path of size v-1. \\
      for each vertex s in V\\
      \ \ \ \ for each vertex t in V, $t \neq s$\\
      \ \ \ \ \ \ \ \ return LongestPath(G, s, t, v-1)
    \end{solution}

    \part \TSP: Given a weighted graph $G$ and bound $B$,
    determine if there is a cycle in $G$ that visits every vertex exactly once
    and has the total sum of edge weights $\leq B$.
    \begin{solution}
      I will show that this problem is a generalization of the Hamiltonian
      Cycle.\\
      Reduction: The Hamiltonian Cycle has input graph G.  Running \TSP on G
      with $B = \infty$ will return true if a cycle is present that visits every
      vertex exactly once, which is by definition a Hamiltonian cycle. This
      shows that the Hamiltonian Cycle reduces to this problem and thus this
      problem is NP-complete. 
    \end{solution}

    \part \prob{Degree\ Restricted\ MST }: Given an undirected graph $G =
    (V,E)$ and an integer $k$, find a spanning tree $T$ of $G$ such that each
    vertex of the tree has maximum degree $k,$ if such a tree exists.
    \begin{solution}
      %Your Solution Goes Here
    \end{solution}

  \end{parts}

  \question (EXAMPLE) \HP: 
  Show that the following problem is NP-complete by reducing from \HC.

  \HP: Given graph $G$, determine if $G$ has a simple {\bf path} (no repeated
  vertices) that visits every vertex exactly once. Here the start and end
  vertices of the path do not need to be the same (or adjacent).
  \begin{solution}
    Note: This solution explains more things than you need to in your
    solutions. This is meant to be instructive.

    \begin{itemize}
      \item To show this problem is in NP, say we are given an input which is
        the graph $G$, as well as the vertices of the path (listed in order).
        Our verifier would check that every vertex was contained in this path,
        and that every proposed edge between adjacent vertices was in the
        graph. If both of these conditions are true, then this is a
        Hamiltonian Path.

      \item 

        Here's our reduction.

        Remember - we are trying to find an algorithm that can efficiently
        solve \HC, a known NP=hard problem, under the assumption that we can
        run an algorithm that can solve \HP, the problem that we are trying to
        show is NP-hard.

        Here's the algorithm:
        Given graph $G$, we're going to construct $G'$ as follows. Pick any
        vertex $v$, and make an extra copy of it called $v'$, and add all the
        same edge connections as $v$.  Now construct another new vertex $s$
        and connect it only to $v$, and another new vertex $t$, and connect it
        only to $v'$. Now, we return $\HP(G')$.

        Here's the proof of correctness of this reduction. We need to show
        that this is CORRECT - that this reduction is an
        algorithm that actually solves \HC. 
        If the input $G$ does have a \HC, we need to return YES. If the input
        $G$ does not have a \HC, we need to return NO.
        
        In this case, that means that showing that $G$ has a \HC if and only
        if $G'$ has an \HP. Here's the proof:

        \begin{itemize}
          \item $G$ has a \HC $\to$ $G'$ has a \HP

            \begin{proof}
              Say $G$ has a \HC. Then since a cycle visits every vertex,
              including $v$, write the vertices of the cycle as $(v, x_1), (x_1,
              x_2), \ldots, (x_{n-1}, v)$. By construction, the edges $(v, x_1),
              (x_1, x_2), \ldots, (x_{n-1}, v')$ are in the graph $G'$, as are
              $(s,v)$ and $(v', t)$. Since the only edges we added were $v', s,
              t$, this new set of edges are a $\HP$ in $G'$.
            \end{proof}

          \item $G'$ has a \HP $\to$ $G$ has a \HC

            \begin{proof}
            Say $G'$ has a \HP.
            Because $s$ and $t$ both have degree 1, any \HP in $G'$ must have
            $s$ and $t$ as its endpoints. Since $s$ only connects to $v$ and
            $t$ to $v'$, then the path must go from $s$ to $v$ to every other
            vertex to $v'$ to $t$. The middle part of this path is a path from
            $v$ to every other vertex back to $v'$ in $G'$. The equivalent
            edges in $G$ would form a path from $v$ to every other vertex and
            back to $v$ again, which are a \HC.
            \end{proof}
        \end{itemize}
    \end{itemize}

    Another valid approach is to add $t$ to a different neighbor of $v$, one
    at a time, until you find a \HP. This would require at most $O(n)$ calls
    to \HP, once for each neighbor of $v$, which is a valid polytime
    reduction.
  \end{solution}

  \question[20] Search \SAT $\to$ Decision \SAT. 
  Describe an algorithm that can {\bf find} a satisfying assignment for a
  given SAT instance (search), if one exists, using an algorithm that can only
  determine {\bf whether or not} a SAT instance \emph{has} a satisfying
  assignment (decision).

  Remember, you may do polynomial work, and call the Decision version of the
  algorithm polynomially many times to solve the Search version.

  Hint: If there is a valid solution to the \SAT instance - use your algorithm
  for Decision \SAT to determine a valid assignment for each variable, one at
  a time.
  \begin{solution}
    Given a decision \SAT algorithm that has as input a list V of variables
    initially set to true and a list C of clauses. This algorithm will determine
    an assignment of values that satisfies the SAT problem. \\

    S-SAT(V, C)\\
    v $\gets$ V.remove(0)\\
    C1 $\gets$ C\\
    for(clause c : C1)\\
    \ \ \ \ c.remove(v) //removes v if it is present\\
    $v \gets false$\\
    $C2 \gets C$\\
    for(clause c : C2)\\
    \ \ \ \ c.remove(v) //removes v if it is present\\

    if(D-SAT(V, C1))\\
    \ \ \ \ return S-SAT(V, C1)\\
    if(D-SAT(V, C2))\\
    \ \ \ \ return S-SAT(V, C2)\\
    return false\\

    This algorithm recursively checks variable assignments one at a time to
    determine an assignment of values that satisfies SAT.  
    %TODO%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    %TODO%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    %TODO%%%%%%%FINISH%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    %TODO%%%%%%%FINISH%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    %TODO%%%%%%%FINISH%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    %TODO%%%%%%%FINISH%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    %TODO%%%%%%%FINISH%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    %TODO%%%%%%%FINISH%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    %TODO%%%%%%%FINISH%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    %TODO%%%%%%%FINISH%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    %TODO%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    %TODO%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    %TODO%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

  \end{solution}

  
  \question[20] \prob{Partition }:
  Show that the following problem is NP-complete by reducing from \SSM.

  Given a list of integers $A$, determine if the integers of $A$ can be
  partitioned into two disjoint lists $A_1, A_2$ that have the same sum.

  For example, the list $A = [7, 3, 2, 9, 6, 13]$ can be partitioned into
  $[7,13]$, $[2,3,6,9]$ each with sum $20$. The list $A = [4,6,8,12]$ cannot be
  partitioned into two equal pieces.
  \begin{solution}
    I will show that \SSM reduces to this problem.\\
    \SSM has an input list A with sum S and a target T. 
    By adding an extra integer to A with a value of S-2T and passing that into
    this algorithm we have it check if the new A with sum S+(S-2T) can be
    partitioned into two disjoint lists each with sum S-T. If such a
    partitioning exists, this means that by removing the added integer we can
    reduce one of the halves to (S-T)-(S-2T), which is T, and the other half is
    equal to S-T, giving us the solution to \SSM and showing that the problem is
    NP-complete. 


    \textbf{\underline{Wrong answer}: Shows reduction in the opposite direction}\\
    This can be done simply by performing \SSM on A with a target of the sum of
    A divided by 2. If this is possible then the subsets not included will also
    have a value of sum A/2. Since a NP-complete problem reduces to this problem
    we know that this problem is also NP-complete. 
  \end{solution}

  \question (Bonus) Consider the following problem.
  You are a hiring manager at a consulting firm.
  You have a set of $n$ available workers $W$ and a set of $m$ tasks $T$ that
  your clients want you to do.
  Each worker $w$ knows how to complete some subset $T[w] \subseteq T$ of the
  tasks, and requires a certain salary to be hired, $S[w]$.
  Each worker, once hired, can complete all the tasks that they know
  how to do. Each completed task earns the company a revenue $R[t]$ per task
  $t$.

  You goal is to maximize your profit, the sum of all revenues earned minus
  the sum of all salaries paid.  Not all tasks have to be assigned; sometimes
  the revenue for completing some tasks may not be worth hiring a worker to do
  it. 

  The decision version of this question asks: Given all this information, $n,
  m, T[w], S[w], R[t]$, can you achieve a profit of at least $k$?  Show that
  this problem is NP-complete.
  \begin{solution}
  \end{solution}
\end{questions}
\end{document}


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: 
